---
title: "Prices Over Space & Time"
subtitle: "Price Analysis: A Fundamental Approach to the Study of Commodity Prices"  

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: TRUE
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(tidyverse)
purduegold <- '#CEB888'
colorscheme <- purduegold
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  #base_color = purduegold,
  primary_color = colorscheme,
  secondary_color = "black",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```



---

# Prices Over Space and Time

**Highlights**

  + Learn the costs of storage for farmers.
  
  + Learn **The Forward Curve** in the futures market.
  
  + Learn the financial calculation of full carry and spread. 
  
  + Learn to interpret the percent of full carry. 
  
  + Learn what drives variation in the basis.
  
**Check Your Understanding**

  + Can you calculate the percent of full carry yourself, given only futures prices and financing costs?
  
---

# In this section we cover how commodity prices behave over time and space

**Following are some important findings:**

  + Commodity futures contracts have an expiration.
  
  + There are always several contracts trading at any given time with maturities.
  
  + Contracts Trading are increasingly farther into the future, and these contracts will eventually expire and no longer be traded.
  
---

# Types Of Contracts:

**There are four different types of contracts:**

  + The contract that is next to expire is called the **Nearby** contract.
  
  
  + The contract that expires next is called the **First Deferred** contract. 
  
  
  + The contract that expires after that is the **Second Deferred** contract. 
  
  
  + The different contracts trading at any given time make up make up what is called **The Forward Curve**, etc. 
  
  
There is valuable information in **The Forward Curve** because it is the market's best guess of what returns to storage will be. 

---

## Storage Costs to the Farmer

**Storage costs include the following:** 

  + **Opportunity cost of money:** 
  
  If they sell at harvest they can use the money for other things.
  
  + **Interest:** 
  
  By deferring the sale of grain, the stockholder may need a bank loan to cover expenses since their main revenue stream is deferred. 
  
  + **Storage fees:** 
  
  Some farmers or stockholders have their own storage space, but many will need to rent storage space.
  
  + **Drying costs:** 
  
  Grain that is just harvested can be around 15% moisture, but must be dried down to closer to 13.5% moisture to safely store for long periods. This involves running a grain dryer that uses fuel and/or electricity.
  
---
## Storage Costs to the Farmer (CONT...)

**Storage costs include the following:**

  + **Shrinkage:** 
  
  When grain is dried, it actually shrinks leaving less bushels to sell after storage. The shrink factor can be 1.25 to 1.4 percent. 
  
  + **Quality deterioration:** 
  
  If the grain is not stored under proper condition, quality can deteriorate, and result in dockage being applied by the buyer at the time of sale. 
  
  + **Cost of handling:** 
  
  Getting the grain into and out of storage results in some costs as well. 
  
---
## Storage Costs to the Farmer (CONT...)


**According to Iowa State University Extension estimated in 2015:**
  
  + The  storing grain until March costs a farmer roughly \$0.45 per bushel.
  
  + The  storing until December cost roughly \$0.30 per bushel which is \$.15 cents per bushel less than storing until March. 
  
  + The he price of the March contract would need to be more than \$0.15 per bushel higher than the December contract.
---

## An Increasing Forward Curve

**Figure 1 illustrates the forward curve on September 26, 2016.**

  + **The Forward Curve** represents the return to storage and it shows the extra money can be made.
  
  + December corn is worth 330 cents per bushel and March corn is worth 340 cents per bushel.
  
  + When stocks are plentiful the market offers a premium to store.

 
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

df <- tibble(xaxis = c("Dec", "Mar", "May", "Jul", "Sep", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(330, 340, 345, 353, 360, 370))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis[order(df$price)]))

g <- ggplot(df, aes(xaxis, price)) + ylim(300,425) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Increasing Forward Curve from Sep 26, 2016", x = "", y = "Price cents per bushel") 
        
       
g
```
  
  
---


+ The price relationship is 'discovered', and changes everyday.

+ When it is upward sloped it is called a "Carry Market" or "Contango Market", or sometimes it is said to be "In Full Carry". 

---

## An Increasing Forward Curve (CONT...)

**This example illustrates a phenomenon that often occurs:**

  + **The Forward Curve** is upward sloped until September and it flattens and returns to storage go away. 
  
  + In September we begin to see the next year's crop come onto the market. 
  
  + In 2015, the market was asking farmers to keep storing through July, but no longer. 
  
  + Holding the grain from July - September could expect to lose as much as 4 cents per month. 
  
---
  
```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

df <- tibble(xaxis = c("Dec", "Mar", "May", "Jul", "Sep", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(390, 400, 408, 412, 405, 410))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(375,415) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Increasing Forward Curve from Sep 25, 2015", x = "", y = "Price cents per bushel") 
        
       
g
```

---

## A Decreasing Forward Curve

  + 2012 was a significant drought year that resulted in: 
  
  1. Poor yields 
  
  2. High prices 
  
  3. Low forecasted ending stocks.
  
  
  + The supplies were tight and **The Forward Curve** tends to be downward sloped. 
  
  
  + The market is incentivising everyone to bring grain onto the market. 
  
---

## A Decreasing Forward Curve (CONT...)

In 04-03-2012, **The Forward Curve** are as shown in figure 3. Its in the spring, before the drought has happened. 


  + The supplies were already tight going into 2012. 
  
  + **The Forward Curve** is downward sloped, sometimes called **Inverted** or **Backward** market. 
  
  + The returns to storage are negative, through the summer of 2012, even before the drought. 
  
  + As of 4-13-2012, the market 'thought' that the 2012 harvest would be good, because the price levels drop substantially. 
  
  + The return to storage between December 2012 and March 2013 is positive on 4-13-2012.
  
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

df <- tibble(xaxis = c("May", "Jul", "Sep", "Dec", "Mar", "May2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(630, 620, 555, 540, 545, 555))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(480,640) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Decreasing Forward Curve from Apr 13, 2012", x = "", y = "Price cents per bushel") 
        
       
g
```
  
---
## A Decreasing Forward Curve (CONT...)

Lets look at the forward curve on 8-01-2012.


  + By August 1, we are in the midst of a major drought, yields will be low along with the ending stocks.

  + In 2013, **The Forward Curve** is downward sloped for the entire marketing year until the next harvest is expected. 
  
  + On 4-13-2012, the market was offering about 5 cents per month to store from December 2012 to March 2013. 
  
  + By 8-01-2012, the market was offering -1 cent for storage during the same time period.

---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

df <- tibble(xaxis = c("Sep", "Dec", "Mar", "May", "Jul", "Sep2", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(800, 800, 800, 799, 797, 698, 610))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(500,900) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Flat Forward Curve from Aug 1, 2012", x = "", y = "Price cents per bushel") 
        
       
g
```

---

## A Decreasing Forward Curve (CONT...)

**To illustrate the forward curve changed between August and December 2012:** 

  + The time in which harvest occurred and we learned the bad yields turned out to be. 
  
  + We show the forward curve on 12-03-2012 in figure 5. 
  
  
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

df <- tibble(xaxis = c("Dec", "Mar", "May", "Jul", "Sep2", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1' ),
             price = c(745, 755, 754, 740, 660, 640))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(500,900) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Flat and Falling Forward Curve from Dec 3, 2012", x = "", y = "Price cents per bushel") 
        
       
g
```


---

## Forward Curve Cases with Hypothetical Data (CONT...)

  + When prices move up or down, the front end of the forward curve is more responsive. 
  
  + We will illustrate this with both increasing prices and decreasing prices. 
  
  + The examples in the next slide show the first five contracts on the forward curve plotted on four consecutive days.
  
  + The price data in these examples are hypothetical, but represent what usually happens to the forward curve when prices.
  
---

## Forward Curve Cases with Hypothetical Data (CONT...)

**Prices Increasing**

+ Figure 6. Forward Curve with Prices Increasing, Contango to backwardation

   + On day 1, the market is clearly in contango, as the forward curve is upward sloped. 
   
   + From day 1 through day 4, the prices are rising each day, with the front end of the forward curve. 
   
   + On day 4, the prices have risen enough that the market is now in backwardation with the front month higher than the first deferred.

---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

xaxis <- c('Dec', 'March', 'May', 'July', 'Sept')
day1 <- c('day 1', 'day 1', 'day 1', 'day 1', 'day 1')
day2 <- c('day 2', 'day 2', 'day 2', 'day 2', 'day 2')
day3 <- c('day 3', 'day 3', 'day 3', 'day 3', 'day 3')
day4 <- c('day 4', 'day 4', 'day 4', 'day 4', 'day 4')
day1p <- as.numeric(c(350, 360, 369, 375, 378))
day2p <- as.numeric(c(355, 364, 372, 377, 379))
day3p <- as.numeric(c(360, 367, 374, 378, 380))
day4p <- as.numeric(c(385, 383, 382, 381, 381))

day1 <- cbind(xaxis, day1, day1p)
day2 <- cbind(xaxis, day2, day2p)
day3 <- cbind(xaxis, day3, day3p)
day4 <- cbind(xaxis, day4, day4p)
df <- rbind(day1, day2, day3, day4)
df <- as.data.frame(df)
colnames(df) <- c('xaxis', 'day', 'price')
# df$xaxis <- factor(df$xaxis, levels = df$xaxis[order(df$price)])
df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis[order(df$price)]))

df$price <- as.numeric(as.character(df$price))
g <- ggplot(df, aes(xaxis, price, colour= day)) + ylim(300,425) + theme_bw() +
        geom_point() + geom_line(aes(group = day)) +
        scale_color_manual(values = c("darkblue", "darkblue", "darkblue", "darkblue")) +
        labs(x = "", y = "Price cents per bushel") + 
        geom_text(aes(label= ifelse(xaxis == "Dec", as.character(day), '')),hjust=1.5, vjust=0) + 
        theme(legend.position = "none")
       
g
```

---

## Forward Curve Cases with Hypothetical Data (CONT...)

**Prices Decreasing**

+ Figure 7. Forward Curve with Prices Decreasing 

  + On day 1, the market is in Contango in this example. 
  
  + From day 1 to day 4, prices are falling. 
  
  + From day 1 to day 3 the forward curve is getting steeper because price declines in Dec are larger than the price declines in March. 
  
  + The price declines in March are larger than the price declines in May. 
  
  + This indicates that from day 1 to day 3, the market is not yet at **Full Carry**; as prices are declining. 
  
  + From day 3 to day 4, however, the market is at **Full Carry** because the price decline is constant all the way up the forward curve. 
  
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

xaxis <- c('Dec', 'March', 'May', 'July', 'Sept')
day1 <- c('day 1', 'day  1', 'day 1', 'day 1', 'day 1')
day2 <- c('day 2', 'day 2', 'day 2', 'day 2', 'day 2')
day3 <- c('day 3', 'day 3', 'day 3', 'day 3', 'day 3')
day4 <- c('day 4', 'day 4', 'day 4', 'day 4', 'day 4')
day1p <- as.numeric(c(350, 360, 369, 375, 378))
day2p <- as.numeric(c(342, 354, 366, 372, 377))
day3p <- as.numeric(c(332, 346, 360, 368, 374))
day4p <- as.numeric(c(327, 341, 355, 363, 369))

day1 <- cbind(xaxis, day1, day1p)
day2 <- cbind(xaxis, day2, day2p)
day3 <- cbind(xaxis, day3, day3p)
day4 <- cbind(xaxis, day4, day4p)
df <- rbind(day1, day2, day3, day4)
df <- as.data.frame(df)
colnames(df) <- c('xaxis', 'day', 'price')
# df$xaxis <- factor(df$xaxis, levels = df$xaxis[order(df$price)])
df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis[order(df$price)]))
df$price <- as.numeric(as.character(df$price))
g <- ggplot(df, aes(xaxis, price, colour= day)) + ylim(300,425) + theme_bw() +
        geom_point() + geom_line(aes(group = day)) +
        scale_color_manual(values = c("darkblue", "darkblue", "darkblue", "darkblue", "darkblue")) +
        labs(x = "", y = "Price cents per bushel") + 
        geom_text(aes(label= ifelse(xaxis == "Dec", as.character(day), '')),hjust=1.5, vjust=0) + 
        theme(legend.position = "none")
       
g
```


---
## Forward Curve Cases with Hypothetical Data (CONT...)


**Some Caveats**

  + The effect of the price changes on the shape of the forward curve is typically observed. 
  
  + There can be fundamental changes in the market that affect the parts of the forward curve. 
  
  + This could cause a larger price change in the middle or back end of the forward curve. 
  
  + The front end of the forward curve will be more volatile than the back end as depicted in figures 6 and 7. 
  
---
## Financial Full Carry

  + The costs of storage to the farmer is the opportunity. 
  
  + The cost of money resulting from deferring a sale to predict with certainty any individual farmer's decision to store. 
  
  + There is a concept called **Financial Full Carry** that simply includes interest costs and the premium charges on shipping certificates.

$$Financial \text{  } Full \text{  } Carry  = ndays(\frac{i}{360}*F + P)$$ 

  + Where $ndays =$ the number of days between the first delivery day in the nearby contract and the first delivery day in the deferred contract. 
  
  + $i =$ three month LIBOR interest rate + 200 basis points.
  
  + $F =$ futures price, and $P =$ the current premium charge on shipping certificates.

---

## Financial Full Carry (CONT...)

+ For example: 

  + There are 90 days between delivery period of the December and March contract. 
  
  + If the LIBOR rate is .3%, financing costs are 200 basis points above LIBOR, the corn futures price is \$3.50 per bushel, and the premium charge on shipping certificates is 0.165 cents per bushel per day, then **Financial Full Carry** is:

  + The CME Group uses simple interest to calculate the financial full carry in other contexts, so we adopt it here for our definition of financial full carry. 

$$Financial \text{  } Full \text{  } Carry = 90*(0.023/360*350 + 0.165 = 16.86 \text{ cents per bushel})$$

---

## Financial Full Carry (CONT...)

  + The **Financial Full Carry** between the December and the March contract would be 16.86 cents. It is called **Financial Full Carry**. 
  
  + The spread between December and March contracts cannot be wider than this amount. 
  
  + If it were wider, say 30 cents, then a storage arbitrage would be possible. 
  
  + Hold the shipping certificate until March 1st at a cost to you of 16.86 cents per bushel. 
  
  + Then, use the shipping certificate to deliver on your short March futures position. 
  
  + Your futures trades just earned 30 cents, while holding the shipping certificate only cost 16.86 cents, leaving you with a profit of 13.14 cents per bushel.  
  
---
## Financial Full Carry (CONT...)


  + The concept of **Financial Full Carry** is really just a benchmark. 
  
  + Most importantly, any individual's ability to capitalize on the arbitrage is predicated at the ability to borrow for 200 basis points over LIBOR. 
  
  + Percent of **Full Financial Carry** is a metric that is widely followed because it gives similar information as the shape of the forward curve across time. 
  
  + In our example, percent of **Full Carry** $=100*30/16.86 = 177.94\%$ (remember this was an extreme example to illustrate the potential for arbitrage). 

  $(\text{Percent of Full Carry } = 100*\frac{Futures \text{  } Calendar \text{  } Spread}{Full \text{  } Financial\text{  } Carry})$ 
  
---

## Calendar Spreads  

  + **The Forward Curve** and returns to storage from the perspective of a farmer or other who holds physical stocks of grain. 
  
  + Speculators watch the price spread between futures contracts and trade them to bet on whether or not returns to storage will increase or decrease. 
  
  + These kinds of spreads are called *Calendar Spreads* and they are done by performing the following type of trade. 
  
---
## Calendar Spreads (CONT...)


+ Following are the same logic about below mentioned:

  + Expected scarcity of stocks 
  
  + Returns to storage 
  
  + Incentives for the market to bring stocks to the market, if the price goes up.
  
  + The nearby contract and front end of the forward curve react the most strongly, deferred contracts will also go up, but by a lesser amount. 
  
  + If the price goes down, the nearby contract will change the most; the deferred contract will also go down, but less so. 
  
---

## A speculator places, the following trades if they are bullish (bearish):

### Bullish - think prices are going up

+ Buy Nearby: Dec 2017

+ Sell Deferred:  March 2018

  + You are betting the prices in general will go up, but the nearby will go up more than the deferred contracts. 
  
  + Any information event suggests supplies will become tighter should make prices go up in general. 
  
  + And, reduce the incentive to the store. Thus, making this a profitable calendar spread trade. 

---
## A speculator places, the following trades if they are bullish (bearish): (CONT...)


### Bearish - think prices are going down

+ Sell Nearby: Dec 2017

+ Buy Deferred: March 2018

  + You are betting thE prices will go down in general, but that the nearby will go down more than the deferred contracts. 
  
  + Any information suggests supplies will become more plentiful should make prices go down in general. 
  
  + And, should increase incentives to store. Thus making the bearish calendar spread profitable. 

---

## Price Variation Over Space

  + Future price represent the expected future price of the commodity in a very specific location. 
  
  + A location that is 'Regular' for delivery is a location designated by the commodity exchange. 
  
  + Where, stocks of a commodity represented by a futures contract may be delivered in fulfillment of the contract. 
  
  + This is where the spot, cash, price, or converge with the futures price.  
---

## Price Variation Over Space (CONT...)


  + The price of the future contract represent the expected future price only at these locations (technically whichever is cheapest to deliver).
  
  + The degree to which the futures price is indicative of the expected future spot price at locations far from Northern Illinois can vary. 

  + In rural U.S, grain elevators, ethanol plants, soybean crushers, feed yards and biodeisel manufacturers dot the landscape every few miles. 
  
  + These entities buy essentially all of the grain and oilseed crop that is not used on-farm for livestock feeding. 
  
  + They post bids to buy every day they are open. They offer to buy as a cash sale, or on forward contract for delivery one to three months ahead.
  
---
  
## Price Variation Over Space (CONT...)

  + In the case of the forward contract, the farmer will go in to the elevator and sign a contract to deliver a specific number of bushels within a specified window of time. 
  
  + Usually, the prices quoted by grain elevators and other prices is relative to the futures contract price, or basis. 
  
  + Depending on how far the location is from the Illinois river, this difference may be large, but still the futures price is the reference point. 
  
  + The basis is often quoted as 'over' or 'under' the futures price. 
  
  + For example, an elevator might post bids to buy for $-27$ cents. This means $27$ cents under the futures price. A bid of $31$ would be read as $31$ cents over the futures price.

---

## Definition of Basis


+ Basis is always defined as Spot Price minus Futures price.

$$Basis = Spot - Futures$$


+ Basis reflects the price differential over space relative to the futures price. Basis is influenced by 

  + Transportation Costs
  
  + Local Supply and Demand Conditions
  
  + Interest and Storage Charges (this reflects that there is also a small time component as well as spatial)
  
  + Other Handling, Shipping and other Costs

---

## Definition of Basis

+ Transportation costs are built into basis because large users of grain are not located in large production region. E.g. 
 
   + Cattle feed yards in Western Kansas and Nebraska
   
   + Chickens in the South
   
   + Hogs in North Carolina 
   
+ Grain is shipped by rail and/or truck to locations across the country. 

+ Areas of grain surplus generally have a negative basis, the spot price is less than the futures. 

+ Areas of grain deficit generally have a positive basis, the spot price is greater than the futures. 

---

## Definition of Basis (CONT...)

  + Local supply and demand conditions are also important. 
  
  + Occasionally, there will be localized production problems. 
  
  + The biggest recent example comes from the demand in ethanol production in the U.S. was felt greatest in Iowa. 
  
  + Billions of gallons of capacity in ethanol production came online in Iowa. 
  
  + The corn basis was affected With additional large consumers of corn located throughout Iowa. 
  
  + There was more localized demand for corn. The ethanol plants and grain elevators had increased localized competition from 2005-10.
  
---

## Terminology

  + Farmers and grain handlers alike watch the basis closely. 
  
  + The discussion of changes in the basis is common. When the basis is increasing, in most cases that means becoming 'less negative'. 
  
  + The basis is **stregthening**. When the basis is decreasing, or becoming 'more negative', we say the basis is **weakening**. 
  
---
class: inverse, center, middle

##CHAPTER END

---