---
title: "Chapter 12 - South American Production" 
subtitle: "Price Analysis: A Fundamental Approach to the Study of Commodity Prices"  

output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: TRUE
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
purduegold <- '#CEB888'
colorscheme <- purduegold
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  #base_color = purduegold,
  primary_color = colorscheme,
  secondary_color = "black",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Fira Mono")
)
```



---

# South American Production


  + Over the last twenty years Brazil and Argentina have grown in prominence on the world stage. 
  
  + They are currently among the most important growers of soybeans and corn, particularly so for soybeans. 

  + Many thanks to Felipe Grimaldi Avileis, my teaching assistant for ACE 427 in 2017. 

  + In Figure 1, we have plotted the production since 1990 of the world's top five producers of corn. 
  
  + Argentina, Brazil, China, the European Union, and the United States round out this group. 
  
  + A clear pattern of increases over time is notable particularly for the U.S., China, and Brazil, generating an upward trend in the production from this group. 

---

## Production of Corn

  + Data for Figures 1-4 come from the [USDA FAS World Production Supply and Demand]historical data-sets for grains and oilseeds.

  + Figure 2 plots the top five countries by exports. 
  
  + We see China and the EU dropping out of the top five with Russia and Ukraine taking their place. 
  
  + While U.S. exports have been mostly flat over this time period, exports from Brazil have been expanding rapidly since about 2004. 
  
  + Next, in table 1, we present the top producers of corn in 2017.
  
  + Table 1: Top 10 World Producers of Corn, 2017
  
---

## Production of Soybeans

  + Now, we turn our attention to the soybean market. 
  
  + In Figure 3, we present production since 1990 from the top five countries. 
  
  + The importance of Brazil and Argentina is much more obvious. 
  
  + While production in the U.S. has a visible upward trend, Brazil and Argentina are really increasing at a rapid rate.
  
---

## Production of Soybeans (CONT...)
  
  + In Figure 4, we plot the exports from the top five countries. 
  
  + From Figure 3 to 4 China and India drop out and are replaced by Canada and Paraguay. 
  
  + The growth in soybean exports from Brazil is quite remarkable in this chart. 
  
  + Table 2 shows the top ten producers of soybeans in the World. Here we see that Paraguay and Bolivia are present in the top ten as well. 
  
---

## South American Production Seasons and Trade Flows

  + South America experiences summer when we in the North America experience winter and vice versa. 
  
  + North American and South American production is nicely complementary in providing corn and soybeans for the export market. 
  
  + U.S. corn and soybean exports are at their highest levels during the 2nd quarter of the marketing year - right after harvest. 
  
  + The planting and growing seasons are underway in South America. 
  
  + The heavy export season from harvesting the previous crop is winding down. 
  
---

## South American Production Seasons and Trade Flows (CONT...)

  + South American farmers are harvesting during our planting and growing seasons (March through June). 

  + The seasonality is a boon for importers because they have the opportunity to buy for seasonally low harvest-time prices.
  
  + The price would increase twice per year rather than just once per year. 

---

### Brazil

  + Brazil, being quite close to the equator, enjoys the ability to double crop most every year. 
  
  + The farmers plant soybeans as the first crop because if they plant soybeans first they get a second crop of corn. 
  
  + If they plant corn first, there is not enough time to get a second crop of soybeans. 
  
  + Yields of first crop corn are higher then yields of second crop corn due to when the rainy season comes. 
  
  + In some places, farmers choose to plant one crop of corn and forego double cropping.  

Figure 6 show production intensity of soybeans by state, and also provides a calendar of when planting and harvest activity of first crop soybeans are taking place. 

---

### Brazil (CONT...)

  + Figure 7 shows the same for second crop corn. 
  
  + While Figure 8 shows the regional production intensity of first crop corn. 
  
  + The highest production of first crop corn is located in regions that have no second crop in Figure 7. 
  
  + In the east, where managing a second crop is more difficult, farmers are more apt to plant a single crop of corn.

---

### Where to find information on Brazilian Crops 

  + Information about Brazillian production the [National Supply Company (Conab)] is the main resource. 
  
  + They publish crop reports monthly, just like the USDA's WASDE report. 
  
  + Typically, the Brazillian report is released a couple of days before the WASDE report. 
  
  + If there is information from Brazil that will affect world stocks or production, these reports will move prices in the U.S. 
  
---

### Where to find information on Brazilian Crops  (CONT...)

  + A calendar of release dates for 2016/2017 and 2017/2018 marketing years is found. 
  
  + The monthly crop reports can be found on the [Conab website]. 
  
  + Ask Google to automatically translate the page to English for you. 
  
  + Navigate through the **Produtos e Servicos** menu, through **Crops**, and choose **Harvesting Safra**. 
  
---

### Where to find information on Brazilian Crops  (CONT...)

  + This brings up a list of reports listed in reverse chronological order by release date. 
  
  + Look for a row of the form **1st Survey - Harvest 2017/2018-Grains**. 
  
  + In the product column will be **Food**. 
  
  + There are three kinds of reports, Food (which contains grains and oilseeds like corn and soybeans), Coffee, and Sugar Cane. 
  
  + Since, the food report is released once per month, the most recent report should always appear on the first page of the table. 
  
---

### Where to find information on Brazilian Crops  (CONT...)


**Area Planted, Yield, and Production**

  + Estimates are given for area planted to soybeans. 
  
  + We have reproduced table 36 from the October 2017 report. 
  
  + Decimals are marked with ',' and 1,000's are marked with '.' - just the opposite of the convention of most English-speaking traditions. 
  
  + We learn that for the country as a whole (bottom row) soybean plantings are up about 2.7% (1.6% - 3.8% low to high range). 
  
  + But, yields are expected to decrease by 8.6%. 
  
  + This leaves productino estimates for the 2017/2018 crop in the range of  -7.1% to -5.1%.
  
---

### Argentina

  + Argentina is more similar to the U.S. in latitude, so they are only able to plant a single crop. 
  
  + Planting of corn begins in September and harvest runs from March until May.
  
  + Planting of soybeans begins in November and harvest runs from April to May. 
  
|Spanish - English Crop Names | |
|:----------------------------|:-------------------|
|Trigo | Wheat  |
|Cebada | Barley  |
|Girasol | Sunflower  |
| Soja | Soybean  |
| Maiz | Corn | 
|Sorgo | Sorghum |

---

### Where to find information on Argentinian Crops


  + The Bolsa de Cereales is the grain exhcange in Argentina. 
  
  + They undertake the responsibility of producing crop reports. 
  
  + The reports most closely followed are the Weekly Agricultural Report, which tells the planting or harvest progress, and gives a general update about market conditions. 
  
  + The other report is the Esdtado Y Condition de Cultivos (ECC), which is released once per month. 
  
  + This has more detailed estimates about planting and growing conditions. 
  
