---
title: "MakeBookAssets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(rnassqs)
library(readxl)
library(lubridate)


# Color Schemes:

purduegold <- '#CEB888'


```

# This file creates assets for my price analysis book. Code can be run to update figures and tables in the book. Assets are located in the `/assets` subdirectory.

# Graphs for `041-Options1.Rmd`

```{r}
prices <- data.frame(price = 400:500)
strike = 450
prem = 10

prices <- prices %>% 
  mutate(callpayoffs = pmax(price - strike, 0) - prem) %>% 
  mutate(putpayoffs = pmax(strike - price, 0) - prem)

longcall <- ggplot(prices, aes(x = price, y = callpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  lims(y = c(-40, 40)) +
  labs(title = "Long Call", x = "", y = "Profit or Loss cents/contract") + 
  theme(plot.title = element_text(hjust = 0.5))

longput <- ggplot(prices, aes(x = price, y = putpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  lims(y = c(-40, 40)) +
  labs(title = "Long Put", x = "", y = "") + 
  theme(plot.title = element_text(hjust = 0.5))

shortcall <- ggplot(prices, aes(x = price, y = -callpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  lims(y = c(-40, 40)) +
  labs(title = "Short Call", x = "Price of Underlying cents/contract", y = "Profit or Loss cents/contract") + 
  theme(plot.title = element_text(hjust = 0.5))

shortput <- ggplot(prices, aes(x = price, y = -putpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  lims(y = c(-40, 40)) +
  labs(title = "Short Put", x = "Price of Underlying cents/contract", y = "") + 
  theme(plot.title = element_text(hjust = 0.5))

figure <- ggarrange(longcall, longput, shortcall, shortput)

figure <- annotate_figure(figure, 
                top = text_grob("Option Profit Diagrams at Expiration", 
                                face = "bold", 
                                size = 14))
ggsave(filename = "Options1-optionplot.png")


```

```{r, fig.width= 6, fig.hight = 4}
prices <- data.frame(price = 1000:1400)
strike = 1200
premc = 47.875
premp = 0.375

prices <- prices %>% 
  mutate(callpayoffs = pmax(price - strike, 0) - premc) %>% 
  mutate(putpayoffs = pmax(strike - price, 0) - premp)

longcall <- ggplot(prices, aes(x = price, y = callpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 1250) +
  lims(y = c(-60, 60)) +
  labs(title = "Long Call", x = "", y = "Profit or Loss cents/contract") + 
  theme(plot.title = element_text(hjust = 0.5))

longput <- ggplot(prices, aes(x = price, y = putpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 1250) +
  lims(y = c(-60, 60)) +
  labs(title = "Long Put", x = "", y = "") + 
  theme(plot.title = element_text(hjust = 0.5))

figure <- ggarrange(longcall, longput)

figure <- annotate_figure(figure, 
                top = text_grob("Option Profit Diagrams at Expiration, 1200 Strikes", 
                                face = "bold", 
                                size = 14))
figure

ggsave(filename = "Options1-optionmoneynessplot.png")

```

```{r, fig.width= 6, fig.hight = 4}
prices <- data.frame(price = 400:500)
strike = 450
premc = 0
premp = 0

prices <- prices %>% 
  mutate(callpayoffs = pmax(price - strike, 0) - premc) %>% 
  mutate(putpayoffs = pmax(strike - price, 0) - premp)

longcall <- ggplot(prices, aes(x = price, y = callpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  lims(y = c(-15, 60)) +
  labs(title = "Long Call", x = "", y = "Profit or Loss cents/contract") + 
  theme(plot.title = element_text(hjust = 0.5))

longput <- ggplot(prices, aes(x = price, y = putpayoffs)) + 
  geom_line(size = 1.5, color = purduegold) +
  theme_bw() + 
  geom_hline(yintercept = 0) + 
  lims(y = c(-15, 60)) +
  labs(title = "Long Put", x = "", y = "") + 
  theme(plot.title = element_text(hjust = 0.5))

figure <- ggarrange(longcall, longput)

figure <- annotate_figure(figure, 
                top = text_grob("Option Payoff Diagrams, 450 Strikes", 
                                face = "bold", 
                                size = 14))
figure

ggsave(filename = "Options1-optionintrinsic.png")

```

# Figures for `042-Options2.RMD`

```{r}
library(gridExtra)
library(OptionPricing)
S <- read.csv(url(paste0("http://ondemand.websol.barchart.com/getHistory.csv?apikey=", KEY, "&symbol=ZS*1&type=weeklyNearest&backAdjust=false&startDate=20100201"))) %>%
   mutate(tradingDay = as.Date(tradingDay)) %>% 
  mutate(sqr_ret = (log(close)- log(lag(close)))^2)

p <- S %>% ggplot(aes(x = tradingDay, y = close)) + 
  geom_line() + 
  theme_bw() +
  labs(x = "", y = "Weekly Closing Price", title = "Soybean Futures")

vol <- S %>% ggplot(aes(x = tradingDay, y = sqr_ret)) + 
  geom_line() + 
  theme_bw() +
  labs(x = "", y = "Squared Returns")

pn <- grid.arrange(p, vol)
ggsave(pn, filename = "Options2-volcluster.png")
```

```{r, fig.width= 6, fig.hight = 4}
prices <- data.frame(price = 400:500)
strike = 450
premc = 0
premp = 0
c <- map(prices$price, BS_EC, T = 0.08, K = strike, r = 0.05, sigma = 0.2) %>% bind_rows()
p <- map(prices$price, BS_EP, T = 0.08, K = strike, r = 0.05, sigma = 0.2) %>% bind_rows()
prices <- prices %>% 
  mutate(callpayoffs = pmax(price - strike, 0) - premc) %>% 
  mutate(putpayoffs = pmax(strike - price, 0) - premp) %>% 
  mutate(callPrice = c$price) %>% 
  mutate(putPrice = p$price)

prices %>% ggplot(aes(x = price, y = callpayoffs)) + 
  geom_line(size = 1.2, color = purduegold) +
  geom_line(size = 1., aes(x = price, y = callPrice)) +
  theme_bw() +
  geom_abline(intercept= -213.9426894, slope= .5, linetype="dashed") +
  geom_abline(intercept= -375.25, slope= .85, linetype="dashed") +
  geom_point(aes(x = 450, y = 11.0573106), size = 2) +
  geom_point(aes(x = 475, y = 28.8629907), size = 2) +
  annotate("segment", x = 445, xend = 449, y = 30, yend = 15, colour = "black", size = 1, arrow = arrow()) +
  annotate("segment", x = 478, xend = 476, y = 10, yend = 22, colour = "black", size = 1, arrow = arrow()) +
  annotate("segment", x = 450, xend = 475, y = 1, yend = 1, colour = "black", size = .5, arrow = arrow()) +
  annotate("text", x = 445, y = 32, label = "delta = .5") +
  annotate("text", x = 478, y = 8, label = "delta = .85") +
  labs(x = "", y = "", title = "Delta of the Option Changes as Asset Price Changes")
ggsave(filename = "Options2-deltahedge.png")
```

# Figures for `044-Options4.Rmd`

```{r, fig.width= 6, fig.hight = 4}
prices <- data.frame(price = 0:500)
strike = 450
strike2 = 425
premc = 0
premp = 17
premp2 = 7.85
c <- map(prices$price, BS_EC, T = 0.33, K = strike, r = 0.05, sigma = 0.2) %>% bind_rows()
p <- map(prices$price, BS_EP, T = 0.33, K = strike, r = 0.05, sigma = 0.2) %>% bind_rows()
p2 <- map(prices$price, BS_EP, T = 0.33, K = strike2, r = 0.05, sigma = 0.2) %>% bind_rows()

prices <- prices %>% 
  mutate(callpayoffs = pmax(price - strike, 0) - premc) %>% 
  mutate(putpayoffs = pmax(strike - price, 0) - premp) %>% 
  mutate(putpayoffs2 = pmax(strike2 - price, 0) - premp2) %>% 
  mutate(callPrice = c$price) %>% 
  mutate(putPrice = p$price) %>% 
  mutate(putPrice2 = p2$price)

prices %>% ggplot(aes(x = price, y = price)) + 
  geom_line(size = 1.2, color = purduegold) +
  theme_bw() +
  labs(x = "Price of Corn", y = "Revanue, $/bu", title = "Farmer Spot Sale per Bushel Revenue")
ggsave(filename = "Options4-spot.png")

prices %>% ggplot(aes(x = price, y = putpayoffs)) + 
  geom_line(size = 1, color = purduegold) +
  geom_line(size = 1., aes(x = price, y = price), color = purduegold) +
  geom_abline(intercept=0, slope= 0, size = .25) +
  theme_bw() +
  labs(x = "Price of Corn", y = "Revanue, $/bu", title = "Farmer Spot Sale and Long Put Revenue per Bushel")
ggsave(filename = "Options4-spotput.png")

prices %>% ggplot(aes(x = price, y = putpayoffs)) + 
  geom_line(size = 1, color = purduegold) +
  geom_line(size = 1., aes(x = price, y = price), color = purduegold) +
  geom_line(size = 1, aes(x = price, y = price + putpayoffs)) +
  geom_abline(intercept=0, slope= 0, size = .25) +
  theme_bw() +
  labs(x = "Price of Corn", y = "Revanue, $/bu", title = "Net Farmer Spot Sale and Long Put Revenue per Bushel")
ggsave(filename = "Options4-spotputcomb.png")

prices %>% ggplot(aes(x = price, y = putpayoffs)) + 
  geom_line(size = 1, color = purduegold) +
  geom_line(size = 1., aes(x = price, y = price), color = purduegold) +
  geom_line(size = 1, aes(x = price, y = price + putpayoffs)) +
  geom_abline(intercept=0, slope= 0, size = .25) +
  theme_bw() +
  lims(x = c(400, 500), y = c(400, 500)) +
  labs(x = "Price of Corn", y = "Revanue, $/bu", title = "Farmer Spot Sale and Long Put Revenue per Bushel")
ggsave(filename = "Options4-spotputcombzoom.png")

prices %>% ggplot(aes(x = price, y = putpayoffs)) + 
  geom_line(size = 1, color = purduegold) +
  geom_line(size = 1., aes(x = price, y = price), color = purduegold) +
  geom_line(size = 1, aes(x = price, y = price + putpayoffs)) +
  geom_line(size = 1, aes(x = price, y = putpayoffs2), linetype = "dashed", color = purduegold) +
  geom_line(size = 1, aes(x = price, y = price + putpayoffs2), linetype = "dashed") +
  geom_abline(intercept=0, slope= 0, size = .25) +
  theme_bw() +
  #lims(x = c(400, 500), y = c(400, 500)) +
  labs(x = "Price of Corn", y = "Revanue, $/bu", title = "Farmer Spot Sale and Long Put Revenue per Bushel")
ggsave(filename = "Options4-spotputcombzoomchoice.png")
```

```{r, fig.width= 6, fig.hight = 8}
prices <- data.frame(price = 0:550)
strike = 450
strike2 = 425
strike3 = 475
premc = 0
premp = 17
premp2 = 7.85
premp3 = 13.5
c <- map(prices$price, BS_EC, T = 0.33, K = strike, r = 0.05, sigma = 0.2) %>% bind_rows()
p <- map(prices$price, BS_EP, T = 0.33, K = strike, r = 0.05, sigma = 0.2) %>% bind_rows()
p2 <- map(prices$price, BS_EP, T = 0.33, K = strike2, r = 0.05, sigma = 0.2) %>% bind_rows()
c3 <- map(prices$price, BS_EC, T = 0.33, K = strike3, r = 0.05, sigma = 0.2) %>% bind_rows()

prices <- prices %>% 
  mutate(callpayoffs = pmax(price - strike, 0) - premc) %>% 
  mutate(putpayoffs = pmax(strike - price, 0) - premp) %>% 
  mutate(putpayoffs2 = pmax(strike2 - price, 0) - premp2) %>% 
  mutate(callpayoffs3 = pmax(price - strike3, 0) - premp3) %>% 
  mutate(callPrice = c$price) %>% 
  mutate(putPrice = p$price) %>% 
  mutate(putPrice2 = p2$price) %>% 
  mutate(callPrice3 = c3$price)



prices %>% ggplot(aes(x = price, y = putpayoffs2)) + 
  geom_line(size = 1, color = purduegold) +
  annotate("segment", x = 100, xend = 100, y = 500, yend = 320, colour = "black", size = .5, arrow = arrow()) +
  annotate("text", x = 105, y = 525, label = "2. Long 425 Put, 7.85 cent prem") +
  geom_line(size = 1., aes(x = price, y = price), color = purduegold) +
  annotate("segment", x = 380, xend = 370, y = 210, yend = 360, colour = "black", size = .5, arrow = arrow()) +
  annotate("text", x = 380, y = 200, label = "1. Naturally Long Spot") +
  geom_line(size = 1., aes(x = price, y = -callpayoffs3), color = purduegold) +
  annotate("segment", x = 215, xend = 295, y = -60, yend = 5, colour = "black", size = .5, arrow = arrow()) +
  annotate("text", x = 200, y = -75, label = "3. Short 475 Call, 13.5 cent prem") +
  geom_line(aes(x = price, y = putpayoffs2 + price - callpayoffs3)) +
  annotate("segment", x = 500, xend = 510, y = 320, yend = 470, colour = "black", size = .5, arrow = arrow()) +
  annotate("text", x = 500, y = 300, label = "4. Net Position") +
  geom_abline(intercept=0, slope= 0, size = .25) +
  theme_bw() +
  labs(x = "", y = "", title = "Farmer Spot Sale and Fence")
ggsave(filename = "Options4-spotfence.png")
```

```{r, fig.width= 6, fig.hight = 4}
prices %>% ggplot(aes(x = price, y = putpayoffs2 + price - callpayoffs3)) + 
  geom_line(size = 1) +
  geom_abline(intercept=0, slope= 0, size = .25) +
  theme_bw() +
  lims(x = c(350, 550), y = c(425, 500)) +
  labs(x = "", y = "", title = "Net Farmer Spot Sale and Fence")
ggsave(filename = "Options4-Netspotfence.png")

```

# Figures for "08-Options3.Rmd"

```{r}
data <- read.csv("cropins_ex - Sheet1.csv") 
  
p1 <- data %>% 
  ggplot(aes(x = index, y = main)) + 
  geom_line(data = subset(data, !is.na(main)), size = 1) + 
  geom_line(data = subset(data, !is.na(Up)), aes(x = index, y = Up), color = "green", size = 1) +
  geom_line(data = subset(data, !is.na(Flat)), aes(x = index, y = Flat), color = "grey", size = 1) +
  geom_line(data = subset(data, !is.na(Down)), aes(x = index, y = Down), color = "red", size = 1) +
  geom_line(data = subset(data, !is.na(BreakEven)), aes(x = index, y = BreakEven), color = "black", size = 1, linetype = "dashed") +
  geom_line(data = subset(data, !is.na(BasePrice)), aes(x = index, y = BasePrice), color = purduegold, size = 1) + 
  theme_bw() + 
  scale_x_continuous(breaks = seq(3,10, 1)) +
  labs(x = "Calendar Month", y = "cents per bushel", title = "Marketing Scenario, Prices Rise Above Breakeven and Base") +
  geom_vline(xintercept = 7) +
  annotate("text", x = 7.18, y = 560, label = "July"  ) +
  annotate("text", x = 8.28, y = 555, label = "Thinking of Locking in High Price"  ) +
  annotate("text", x = 10.15, y = 560, label = "560" ) +
  annotate("text", x = 10.15, y = 535, label = "535" ) +
  annotate("text", x = 10.15, y = 510, label = "510" ) +
  annotate("text", x = 9.85, y = 500, label = "Base 500") +
  annotate("text", x = 9.85, y = 490, label = "BreakEven 490") 
  
ggsave(p1, filename = "cropins_ex.png", width = 8)






```

# Figures for "PrimerforGrain.rmd"

```{r}
key = 'BD652A51-1B61-3442-B80A-9B4030E2F382'


# apikey <- "YOUR API KEY"
nassqs_auth(key)

params1 <- list(commodity_desc = c("CORN", "SOYBEANS", "WHEAT"), 
                year__GE = 1926, 
                agg_level_desc = "NATIONAL", 
                statisticcat_desc = c("YIELD", "Area Planted", "Production", "PRICE RECEIVED"),
                prodn_practice_desc = "ALL PRODUCTION PRACTICES",
                util_practice_desc = c("GRAIN", "ALL UTILIZATION PRACTICES"),
                reference_period_desc = c("YEAR", "Marketing Year"),
                freq_desc = "ANNUAL"
                #prodn_practice_desc = "NON-IRRIGATED"
                ) # GE means "greater than or equal to"
data1   <- nassqs(params1)
data1$Value <- as.numeric(gsub(",", "", as.character(data1$Value)))


corn1 <- data1 %>% 
  select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
  filter( commodity_desc == "CORN") %>%
  filter(statisticcat_desc == "AREA PLANTED") %>% 
   mutate(year = as.numeric(year)) %>%
  ggplot(aes(x = year, y = Value/1000000)) + 
  geom_col(width = .7) +
  geom_line(data = data1 %>% 
  select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
  filter( commodity_desc == "CORN") %>%
  filter(statisticcat_desc == "YIELD") %>%
   mutate(year = as.numeric(year)), aes(x = year, y = Value), color = purduegold, size = 1.5) +
  theme_bw() +
  lims(y = c(0, 200)) +
  labs(title = "U.S. Corn Acreage and Yield", x = "", y = "Millions of Acres, and Bu/Acre, respectively") + 
  scale_x_continuous(n.breaks = 25)

corn1  
ggsave(corn1, filename = "PrimerforGrain_CornAcandY.png")



corn2 <- data1 %>% 
  filter( commodity_desc == "CORN") %>%
  filter(statisticcat_desc == "PRODUCTION") %>%
  filter(short_desc == "CORN, GRAIN - PRODUCTION, MEASURED IN BU") %>%
  filter( domaincat_desc == "NOT SPECIFIED") %>% 
  filter(source_desc == "SURVEY") %>% 
   mutate(year = as.numeric(year)) %>%
  select(c(commodity_desc, statisticcat_desc, short_desc, domaincat_desc, source_desc, year, Value )) %>% 
  ggplot(aes(x = year, y = Value/1000000000)) + 
  geom_col(width = 1.0) +
  geom_line(data = data1 %>% 
  filter( commodity_desc == "CORN") %>%
  filter(statisticcat_desc == "PRICE RECEIVED") %>%
  filter(reference_period_desc == "MARKETING YEAR") %>% 
  filter(short_desc == "CORN, GRAIN - PRICE RECEIVED, MEASURED IN $ / BU") %>% 
    
   mutate(year = as.numeric(year)) %>%
 select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
   mutate(year = as.numeric(year)), aes(x = year, y = Value), color = purduegold, size = 1.5) +
  theme_bw() +
  lims(y = c(0, 16)) +
  labs(title = "U.S. Corn Production and Prices", x = "", y = "Billions of Bushels, and $/bu, respectively") + 
  scale_x_continuous(n.breaks = 25)
  
#+
 # scale_y_continuous(
#    name = "$/bu", 
 #   sec.axis = sec_axis(~ . * 1.5  , name = "Billions of Bushels"), 
  #  limits = c(0, 17)) 

corn2
ggsave(corn2, filename = "PrimerforGrain_CornProdand$.png")

```

```{r}
soy1 <- data1 %>% 
  select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
  filter( commodity_desc == "SOYBEANS") %>%
  filter(statisticcat_desc == "AREA PLANTED") %>% 
  mutate(year = as.numeric(year)) %>%
  ggplot(aes(x = year, y = Value/1000000)) + 
  geom_col(width = 1) +
  geom_line(data = data1 %>% 
  select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
  filter( commodity_desc == "SOYBEANS") %>%
  filter(statisticcat_desc == "YIELD") %>%
   mutate(year = as.numeric(year)), aes(x = year, y = Value), color = purduegold, size = 1.5)  +
  theme_bw() +
  lims(y = c(0, 100)) +
  labs(title = "U.S. Soybean Acreage and Yield", x = "", y = "Millions of Acres, and Bu/Acre, respectively") + 
  scale_x_continuous(n.breaks = 25)

soy1  
ggsave(soy1, filename = "PrimerforGrain_SoyAcandY.png")



soy2 <- data1 %>% 
  filter( commodity_desc == "SOYBEANS") %>%
  filter(statisticcat_desc == "PRODUCTION") %>%
  filter(short_desc == "SOYBEANS - PRODUCTION, MEASURED IN BU") %>%
  filter( domaincat_desc == "NOT SPECIFIED") %>% 
  filter(source_desc == "SURVEY") %>% 
   mutate(year = as.numeric(year))%>%
  select(c(commodity_desc, statisticcat_desc, short_desc, domaincat_desc, source_desc, year, Value )) %>% 
  ggplot(aes(x = year, y = Value/1000000000)) + 
  geom_col(width = 1.0) +
  geom_line(data = data1 %>% 
  filter( commodity_desc == "SOYBEANS") %>%
  filter(statisticcat_desc == "PRICE RECEIVED") %>%
  filter(reference_period_desc == "MARKETING YEAR") %>% 
  filter(short_desc == "SOYBEANS - PRICE RECEIVED, MEASURED IN $ / BU") %>% 
 select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
   mutate(year = as.numeric(year)), aes(x = year, y = Value), color = purduegold, size = 1.5) +
  theme_bw() +
  lims(y = c(0, 16)) +
  labs(title = "U.S. Soybean Production and Prices", x = "", y = "Billions of Bushels, and $/bu, respectively") + 
  scale_x_continuous(n.breaks = 25)
  
#+
 # scale_y_continuous(
#    name = "$/bu", 
 #   sec.axis = sec_axis(~ . * 1.5  , name = "Billions of Bushels"), 
  #  limits = c(0, 17)) 

soy2
ggsave(soy2, filename = "PrimerforGrain_SoyProdand$.png")


```

```{r}
wheat1 <- data1 %>%  
  select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
  filter( commodity_desc == "WHEAT") %>%
  filter(short_desc == "WHEAT - ACRES PLANTED") %>%  
  filter(statisticcat_desc == "AREA PLANTED") %>% 
   mutate(year = as.numeric(year)) %>%
  ggplot(aes(x = year, y = Value/1000000)) + 
  geom_col(width = 1) +
  geom_line(data = data1 %>% 
  select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
  filter( commodity_desc == "WHEAT") %>%
  filter(short_desc == "WHEAT - YIELD, MEASURED IN BU / ACRE") %>%
  filter(statisticcat_desc == "YIELD") %>%
   mutate(year = as.numeric(year)), aes(x = year, y = Value) , color = purduegold, size = 1.5) +
  theme_bw() +
  lims(y = c(0, 100)) +
  labs(title = "U.S. Wheat Acreage and Yield", x = "", y = "Millions of Acres, and Bu/Acre, respectively") + 
  scale_x_continuous(n.breaks = 25)

wheat1  
ggsave(wheat1, filename = "PrimerforGrain_WheatAcandY.png")



wheat2 <- data1 %>% 
  filter( commodity_desc == "WHEAT") %>%
  filter(statisticcat_desc == "PRODUCTION") %>%
  filter(short_desc == "WHEAT - PRODUCTION, MEASURED IN BU") %>%
  filter( domaincat_desc == "NOT SPECIFIED") %>% 
  filter(source_desc == "SURVEY") %>% 
   mutate(year = as.numeric(year)) %>%
  select(c(commodity_desc, statisticcat_desc, short_desc, domaincat_desc, source_desc, year, Value )) %>% 
  ggplot(aes(x = year, y = Value/1000000000)) + 
  geom_col(width = 1.0) +
  geom_line(data = data1 %>% 
  filter( commodity_desc == "WHEAT") %>%
  filter(statisticcat_desc == "PRICE RECEIVED") %>%
  filter(reference_period_desc == "MARKETING YEAR") %>% 
  filter(short_desc == "WHEAT - PRICE RECEIVED, MEASURED IN $ / BU") %>% 
 select(c(commodity_desc, statisticcat_desc, short_desc, year, Value )) %>%
   mutate(year = as.numeric(year)), aes(x = year, y = Value),  color = purduegold, size = 1.5) +
  theme_bw() +
  lims(y = c(0,9)) +
  labs(title = "U.S. Wheat Production and Prices", x = "", y = "Billions of Bushels, and $/bu, respectively") + 
  scale_x_continuous(n.breaks = 25)
  
#+
 # scale_y_continuous(
#    name = "$/bu", 
 #   sec.axis = sec_axis(~ . * 1.5  , name = "Billions of Bushels"), 
  #  limits = c(0, 17)) 

wheat2
ggsave(wheat2, filename = "PrimerforGrain_WheatProdand$.png")


```

```{r}
library(tidyverse)
library(readxl)
library(httr)
library(tidyr)


tmp <- tempfile(fileext = ".xls")

#USDA data discontinued for now. 
# httr::GET(url = "https://www.ers.usda.gov/webdocs/DataFiles/50048/Feed%20Grains%20Yearbook%20Tables-All%20Years.xls",
#           write_disk( tmp, overwrite = TRUE) )


tmp = 'C:/Users/mindy/Desktop/Documents/PriceAnalysis/PriceAnalysis/assets/feed-grains-yearbook-historical.xlsx'

data_xl <- read_excel(tmp, sheet = "FGYearbookTable04", 
                      skip = 4, 
                      col_names = c("MktYear", "Quarter", "BeginningStocks", "Production", "Imports", 
                                    "TotalSupply", "FoodAlcoholInd", "Seed", "FeedandResidual", 
                                    "TotalDomestic", "Exports", "TotalUse", "EndingStocks")) %>%
  fill(MktYear) %>% 
  # Make columns for more standard date handling
  mutate(month = ifelse(Quarter =="Q1 September-November", "09-01", 
                        ifelse(Quarter == "Q2 December-February", "12-01", 
                               ifelse(Quarter == "Q3 March-May", "03-01", 
                                      ifelse(Quarter == "Q4 June-August", "06-01",
                                             ifelse(Quarter == "MY September-August", "09-01", NA)))))) %>%
  mutate(year = ifelse(Quarter =="Q1 September-November" | Quarter == "Q2 December-February" | Quarter == "MY  September-August", as.numeric(str_sub(MktYear, 1,4)), 
                       as.numeric(str_sub(MktYear, 1,4)) +1 )) %>%
  filter(!is.na(month) & !is.na(year)) %>%
  mutate(date = as.Date(paste0(as.character(year), "-", as.character(month)))) # Now we have a proper date column.

# Make tidy getting ready for ggplot
data_tidy <- data_xl %>% 
  select(c(-MktYear, -month, -year)) %>%
  pivot_longer(c(-date, -Quarter), names_to = "UseCategory", values_to = "Value")

data_tidy

# Plot the Marketing Year Data
CornUse <- data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("FoodAlcoholInd", "Seed", "FeedandResidual", "Exports", "EndingStocks")) %>%
  ggplot(aes(x = date, y = Value, fill = UseCategory)) + 
  geom_area() + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Use Categories Since 1975")


CornUse
ggsave(CornUse, filename = "PrimerforGrain_CornUse.png")

```

```{r}
tmp1 <- tempfile(fileext = ".xlsx")

# httr::GET(url = "https://www.ers.usda.gov/webdocs/DataFiles/52218/OilCropsAllTables.xlsx?v=486",
#           write_disk( tmp1) )

# have to get manually for now
# "https://www.ers.usda.gov/data-products/oil-crops-yearbook"
tmp1 = 'C:/Users/mindy/Desktop/Documents/PriceAnalysis/PriceAnalysis/assets/OilCropsAllTables.xlsx'
data_xl <- read_excel(tmp1, sheet = "tab3", 
                     skip = 6, 
                      col_names = c("MktYear", "BeginningStocks", "Production", "Imports", 
                                    "TotalSupply", "Crush", "Exports", "SeedFeedandResidual", 
                                    "Total", "EndingStocks", "Prices")) 

# Make tidy getting ready for ggplot
data_xl$Prices <- as.numeric(data_xl$Prices)
data_xl <- data_xl[1:43,]  # Increase this by one to update. Have to cut off junk at bottom
data_xl$MktYear <- str_sub(data_xl$MktYear,1,4) 
data_xl$MktYear <- as.numeric(data_xl$MktYear)
data_tidy <- data_xl %>% 
  pivot_longer(c(-MktYear), names_to = "UseCategory", values_to = "Value")


data_tidy

# Plot the Marketing Year Data
SoybeanUse <- data_tidy %>% 
  filter(UseCategory %in% c("Crush", "Exports", "SeedFeedandResidual", "EndingStocks")) %>%
  ggplot(aes(x = MktYear, y = Value, fill = UseCategory)) + 
  geom_area() + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Soybean Use Categories Since 1980") +
   scale_x_continuous(n.breaks = 10)

SoybeanUse
ggsave(SoybeanUse, filename = "PrimerforGrain_SoyUse.png")


```

```{r}
CornPrices<- data1 %>% 
  filter( commodity_desc == "CORN") %>%
  filter(statisticcat_desc == "PRICE RECEIVED") %>%
  filter(reference_period_desc == "MARKETING YEAR") %>% 
  filter(short_desc == "CORN, GRAIN - PRICE RECEIVED, MEASURED IN $ / BU") %>% 
 select(c(commodity_desc, statisticcat_desc, short_desc, year, Value ))%>%
   mutate(year = as.numeric(year)) %>% 
  ggplot(aes(x = year, y = Value)) + 
  geom_line(color = purduegold, size = 1.5) +
  theme_bw() +
  lims(y = c(0, 8)) +
  labs(title = "U.S. Corn Prices", x = "", y = "$/bu") + 
  scale_x_continuous(n.breaks = 25)

CornPrices
ggsave(CornPrices, filename = "PrimerforGrain_CornPrices.png")
```

# Figures for `05-PricesSpaceTime.Rmd`

```{r, echo = FALSE, message = FALSE, warning=FALSE}
library(tidyr)
library(ggplot2)

df <- tibble(xaxis = c("Dec", "Mar", "May", "Jul", "Sep", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(330, 340, 345, 353, 360, 370))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis[order(df$price)]))

g <- ggplot(df, aes(xaxis, price)) + ylim(300,425) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Increasing Forward Curve from Sep 26, 2016", x = "", y = "Price cents per bushel") 
        
g       
ggsave(filename = "PricesSpaceTime-increasing-9-26-2016.png")



df <- tibble(xaxis = c("Dec", "Mar", "May", "Jul", "Sep", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(390, 400, 408, 412, 405, 410))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(375,415) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Increasing Forward Curve from Sep 25, 2015", x = "", y = "Price cents per bushel") 
        
       
g
ggsave(filename = "PricesSpaceTime-increasing-9-25-2015.png")


df <- tibble(xaxis = c("May", "Jul", "Sep", "Dec", "Mar", "May2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(630, 620, 555, 540, 545, 555))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(480,640) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Decreasing Forward Curve from Apr 13, 2012", x = "", y = "Price cents per bushel") 
        
       
g
ggsave(filename = "PricesSpaceTime-decreasing-4-13-2012.png")



df <- tibble(xaxis = c("Sep", "Dec", "Mar", "May", "Jul", "Sep2", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1'),
             price = c(800, 800, 800, 799, 797, 698, 610))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(500,900) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Flat Forward Curve from Aug 1, 2012", x = "", y = "Price cents per bushel") 
        
       
g
ggsave(filename = "PricesSpaceTime-flat-8-01-2012.png")


df <- tibble(xaxis = c("Dec", "Mar", "May", "Jul", "Sep2", "Dec2"),
             day = c('day 1', 'day 1', 'day 1', 'day 1', 'day 1', 'day 1' ),
             price = c(745, 755, 754, 740, 660, 640))

df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis))

g <- ggplot(df, aes(xaxis, price)) + ylim(500,900) + theme_bw() +
        geom_point() + 
        geom_line(aes(group = day)) +
        scale_color_manual(values = c(purduegold, purduegold, purduegold, purduegold, purduegold, purduegold)) +
        labs(title = "Flat and Falling Forward Curve from Dec 3, 2012", x = "", y = "Price cents per bushel") 
        
       
g
ggsave(filename = "PricesSpaceTime-flat-12-03-2012.png")

xaxis <- c('Dec', 'March', 'May', 'July', 'Sept')
day1 <- c('day 1', 'day 1', 'day 1', 'day 1', 'day 1')
day2 <- c('day 2', 'day 2', 'day 2', 'day 2', 'day 2')
day3 <- c('day 3', 'day 3', 'day 3', 'day 3', 'day 3')
day4 <- c('day 4', 'day 4', 'day 4', 'day 4', 'day 4')
day1p <- as.numeric(c(350, 360, 369, 375, 378))
day2p <- as.numeric(c(355, 364, 372, 377, 379))
day3p <- as.numeric(c(360, 367, 374, 378, 380))
day4p <- as.numeric(c(385, 383, 382, 381, 381))

day1 <- cbind(xaxis, day1, day1p)
day2 <- cbind(xaxis, day2, day2p)
day3 <- cbind(xaxis, day3, day3p)
day4 <- cbind(xaxis, day4, day4p)
df <- rbind(day1, day2, day3, day4)
df <- as.data.frame(df)
colnames(df) <- c('xaxis', 'day', 'price')
# df$xaxis <- factor(df$xaxis, levels = df$xaxis[order(df$price)])
df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis[order(df$price)]))

df$price <- as.numeric(as.character(df$price))
g <- ggplot(df, aes(xaxis, price, colour= day)) + ylim(300,425) + theme_bw() +
        geom_point() + geom_line(aes(group = day)) +
        scale_color_manual(values = c("darkblue", "darkblue", "darkblue", "darkblue")) +
        labs(x = "", y = "Price cents per bushel") + 
        geom_text(aes(label= ifelse(xaxis == "Dec", as.character(day), '')),hjust=1.5, vjust=0) + 
        theme(legend.position = "none")
       
g
ggsave(filename = "PricesSpaceTime-pricesincreasing-hyp.png")


xaxis <- c('Dec', 'March', 'May', 'July', 'Sept')
day1 <- c('day 1', 'day  1', 'day 1', 'day 1', 'day 1')
day2 <- c('day 2', 'day 2', 'day 2', 'day 2', 'day 2')
day3 <- c('day 3', 'day 3', 'day 3', 'day 3', 'day 3')
day4 <- c('day 4', 'day 4', 'day 4', 'day 4', 'day 4')
day1p <- as.numeric(c(350, 360, 369, 375, 378))
day2p <- as.numeric(c(342, 354, 366, 372, 377))
day3p <- as.numeric(c(332, 346, 360, 368, 374))
day4p <- as.numeric(c(327, 341, 355, 363, 369))

day1 <- cbind(xaxis, day1, day1p)
day2 <- cbind(xaxis, day2, day2p)
day3 <- cbind(xaxis, day3, day3p)
day4 <- cbind(xaxis, day4, day4p)
df <- rbind(day1, day2, day3, day4)
df <- as.data.frame(df)
colnames(df) <- c('xaxis', 'day', 'price')
# df$xaxis <- factor(df$xaxis, levels = df$xaxis[order(df$price)])
df$xaxis <- factor(df$xaxis, levels = unique(df$xaxis[order(df$price)]))
df$price <- as.numeric(as.character(df$price))
g <- ggplot(df, aes(xaxis, price, colour= day)) + ylim(300,425) + theme_bw() +
        geom_point() + geom_line(aes(group = day)) +
        scale_color_manual(values = c("darkblue", "darkblue", "darkblue", "darkblue", "darkblue")) +
        labs(x = "", y = "Price cents per bushel") + 
        geom_text(aes(label= ifelse(xaxis == "Dec", as.character(day), '')),hjust=1.5, vjust=0) + 
        theme(legend.position = "none")
       
g
ggsave(filename = "PricesSpaceTime-pricesdecreasing-hyp.png")

```

#2023 updated to here

# Figures for `08-ForecastingProduction.Rmd`

```{r, fig.width = 8, fig.height= 5}
nassqs_auth(key)

params1 <- list(commodity_desc = c("CORN", "SOYBEANS", "WHEAT"), 
                year__GE = 1800, 
                agg_level_desc = "NATIONAL", 
                statisticcat_desc = c("YIELD", "Area Planted", "Area Harvested", "Production", "PRICE RECEIVED"),
                prodn_practice_desc = "ALL PRODUCTION PRACTICES",
                util_practice_desc = c("GRAIN", "ALL UTILIZATION PRACTICES"),
                reference_period_desc = c("YEAR", "Marketing Year"),
                freq_desc = "ANNUAL"
                #prodn_practice_desc = "NON-IRRIGATED"
                ) # GE means "greater than or equal to"
data1   <- nassqs(params1)
data1$Value <- as.numeric(gsub(",", "", as.character(data1$Value)))
data1

corn2 <- data1 %>% 
  filter( commodity_desc == "CORN" | commodity_desc == "SOYBEANS") %>%
  filter(statisticcat_desc == "AREA PLANTED") %>%
  #filter(short_desc == "CORN, GRAIN - PRODUCTION, MEASURED IN BU") %>%
  filter( domaincat_desc == "NOT SPECIFIED") %>% 
  filter(source_desc == "SURVEY") %>%
  select(c(commodity_desc, statisticcat_desc, short_desc, domaincat_desc, source_desc, year, Value )) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(x = year, y = Value/1000000, color = commodity_desc)) + 
  geom_line() +
  geom_line(data = data1 %>% 
    filter( commodity_desc == "CORN" | commodity_desc == "SOYBEANS") %>%
    filter(statisticcat_desc == "PRICE RECEIVED") %>%
    filter(reference_period_desc == "MARKETING YEAR") %>% 
    mutate(year = as.numeric(year)) %>% 
    filter(year> 1910) %>% 
    select(c(commodity_desc, year, Value )) %>% 
    pivot_wider(names_from = "commodity_desc", values_from = "Value", values_fn = {mean}) %>% 
    mutate(CornSoyRatio = lead(CORN, 2)/lead(SOYBEANS, 2)*100), 
     aes(x = year, y = CornSoyRatio, color = "Last-Year-Corn/Soy-PriceRatio")) +
  theme_bw() +
  #lims(y = c(0, 16)) +
  labs(title = "U.S. Corn and Soybeans Planted Acres", x = "", y = "Millions of Acres") + 
  scale_x_continuous(n.breaks = 15)
  
corn2

ggsave(filename = "ForecastingProduction-PlantedAcres.png")
```

```{r, fig.width = 6, fig.height= 5}

corn <- read.csv("zcz17.csv") %>% select( timestamp, close)
soybeans <- read.csv("zsx17.csv") %>% select( timestamp, close)
       
temp <- inner_join(corn, soybeans, by = "timestamp")
colnames(temp) <- c("timestamp", "Corn", "Soybeans")

 
temp %>% mutate(CornSoyRatio = Corn/Soybeans*100) %>%
  mutate(timestamp = as.Date(timestamp)) %>% 
  filter(timestamp < as.Date("2017-07-01")) %>%
  ggplot(aes(x = timestamp, y = CornSoyRatio)) +
   geom_line() +
  theme_bw() +
  #lims(y = c(0, 16)) +
  labs(title = "2017 Harvest Price Ratio Corn and Soybeans", x = "", y = "Corn/Soy*100")

ggsave(filename = "ForecastingProduction-PriceRatio.png")


corn2 <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "AREA PLANTED" | statisticcat_desc == "AREA HARVESTED") %>%
  filter( domaincat_desc == "NOT SPECIFIED") %>% 
  filter(source_desc == "SURVEY") %>%
  select(c(year, commodity_desc, statisticcat_desc, Value )) %>% 
  mutate(year = as.numeric(year)) %>% 
  filter(year > 1998) %>% 
  pivot_wider(names_from = 'statisticcat_desc', values_from = "Value") %>%
  mutate(DIFFERENCE = `AREA PLANTED` - `AREA HARVESTED`) %>% 
  pivot_longer(cols = c("AREA PLANTED", "AREA HARVESTED", "DIFFERENCE"), names_to = "ACRES") %>%
  ggplot(aes(x = year, y = value/1000000, color = ACRES)) + 
  geom_line() +
  theme_bw() +
  scale_x_continuous(n.breaks = 5) +
  #lims(y = c(0, 16)) +
  labs(title = "Corn Planted Acres and Harvested Acres Since 1999", x = "", y = "Millions of Acres")

corn2
ggsave(filename = "ForecastingProduction-PAHA.png")



corn2 <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "YIELD") %>%
  select(c(year, commodity_desc, statisticcat_desc, Value )) %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(x = year, y = Value)) + 
  geom_line(color = "black", size = .75) +
  geom_smooth(method = "lm", color = purduegold) +
  theme_bw() +
  scale_x_continuous(n.breaks = 10) +
  #lims(y = c(0, 16)) +
  labs(title = "Corn Yields, bu/acre", x = "", y = "")

corn2
ggsave("ForecastProduction-yieldsall.png")

corn2 <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "YIELD") %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1950) %>% 
  select(c(year, commodity_desc, statisticcat_desc, Value )) %>% 
  ggplot(aes(x = year, y = Value)) + 
  geom_line(color = "black", size = .75) +
  geom_smooth(method = "lm", color = purduegold) +
  theme_bw() +
  scale_x_continuous(n.breaks = 10) +
  #lims(y = c(0, 16)) +
  labs(title = "Corn Yields 1950-present, bu/acre", x = "", y = "")

corn2
ggsave("ForecastProduction-yields1950.png")


corn2 <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "YIELD") %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1980) %>% 
  select(c(year, commodity_desc, statisticcat_desc, Value )) %>% 
  ggplot(aes(x = year, y = Value)) + 
  geom_line(color = "black", size = .75) +
  geom_smooth(method = "lm", color = purduegold) +
  theme_bw() +
  scale_x_continuous(n.breaks = 10) +
  #lims(y = c(0, 16)) +
  labs(title = "Corn Yields 1980-present, bu/acre", x = "", y = "")

corn2
ggsave("ForecastProduction-yields1980.png")


temp1 <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "YIELD") %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1980) %>% 
  select(c(year, commodity_desc, statisticcat_desc, Value ))
  
temp1 <- lm(Value ~ year, temp1)

temp <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "YIELD") %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1950) %>% 
  select(c(year, commodity_desc, statisticcat_desc, Value ))
  
temp <- lm(Value ~ year, temp)


corn2 <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "YIELD") %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1980) %>% 
  select(c(year, commodity_desc, statisticcat_desc, Value )) %>% 
  mutate(LevelDeviation = Value - (temp1$coefficients[1] + temp1$coefficients[2]*year)) %>% 
  ggplot(aes(x = year, y = LevelDeviation)) + 
  geom_line(color = purduegold, size = .75) +
  theme_bw() +
  scale_x_continuous(n.breaks = 10) +
  #lims(y = c(0, 16)) +
  labs(title = "Deviations from Trend Yield (1980)", x = "", y = "")

corn2
ggsave("ForecastProduction-LevelDeviation.png")

corn2 <- data1 %>% 
  filter( commodity_desc == "CORN" ) %>%
  filter(statisticcat_desc == "YIELD") %>%
  mutate(year = as.numeric(year)) %>% 
  filter(year >= 1980) %>% 
  select(c(year, commodity_desc, statisticcat_desc, Value )) %>% 
  mutate(LevelDeviation = 100*(log(Value) - log((temp1$coefficients[1] + temp1$coefficients[2]*year)))) %>% 
  ggplot(aes(x = year, y = LevelDeviation)) + 
  geom_line(color = purduegold, size = .75) +
  theme_bw() +
  scale_x_continuous(n.breaks = 10) +
  #lims(y = c(0, 16)) +
  labs(title = "% Deviations from Trend Yield (1980)", x = "", y = "")

corn2
ggsave("ForecastProduction-PercDeviation.png")


# Getting crop condition ratings

params1 <- list(commodity_desc = c("CORN"), 
                year__GE = 1986, 
                agg_level_desc = "NATIONAL", 
                statisticcat_desc = c("YIELD", "CONDITION"),
                prodn_practice_desc = "ALL PRODUCTION PRACTICES",
                util_practice_desc = c("GRAIN", "ALL UTILIZATION PRACTICES"),
                reference_period_desc = c("YEAR", "WEEK #30", "WEEK #31", "WEEK #32", "WEEK #33", "WEEK #34", "WEEK #35", "WEEK #36", "WEEK #37",
                                          "WEEK #38", "WEEK #39", "WEEK #40", "WEEK #41", "WEEK #42", "WEEK #43", "WEEK #44", "WEEK #45", 
                                          "WEEK #46"),
                freq_desc = c("ANNUAL", "WEEKLY")
                #prodn_practice_desc = "NON-IRRIGATED"
                ) # GE means "greater than or equal to"
data1   <- nassqs(params1)
data1$Value <- as.numeric(gsub(",", "", as.character(data1$Value)))

data1 <- data1 %>% filter(str_detect(short_desc, "EXCELLENT") | str_detect(short_desc, "GOOD") | str_detect(short_desc, "YIELD")) %>%
  select(c(year, week_ending, short_desc, Value)) %>% 
  mutate(type = case_when(str_detect(short_desc, "EXCELLENT") ~ "Condition",
                          str_detect(short_desc, "GOOD") ~ "Condition",
                          TRUE ~ "YIELD"))
  

dataGE <- data1 %>% filter(type == "Condition") %>% 
  group_by(year) %>%
  pivot_wider(names_from = short_desc, values_from = Value) %>%
  slice(n()) %>% 
  mutate(`% Good + Excellent` = `CORN - CONDITION, MEASURED IN PCT EXCELLENT` + `CORN - CONDITION, MEASURED IN PCT GOOD`) %>%
  select(year, `% Good + Excellent`) %>% 
  ungroup()

dataY <- data1 %>% filter(type == "YIELD") %>% 
  select(year, Value)

colnames(dataY) <- c("year", "Yield")

temp1 <- lm(Yield ~ year, dataY)

inner_join(dataGE, dataY, by = "year") %>%
   mutate(year = as.numeric(year)) %>% 
  mutate(`Percent Deviation from Trend` = 100*(log(Yield) - log((temp1$coefficients[1] + temp1$coefficients[2]*year)))) %>% 
  pivot_longer(cols = c(`% Good + Excellent`, `Percent Deviation from Trend`), names_to = "Series", values_to = "Value") %>% 
  ggplot(aes(x = year, y = Value, color = Series)) + 
  geom_line(size = .75) + 
  theme_bw() + 
  scale_color_manual(values=c("black", purduegold)) + 
  labs(title = "End of Season G + E Crop Condition Ratings and Final Yield")

ggsave("ForecastProduction-PercDeviationVSGE.png")


inner_join(dataGE, dataY, by = "year") %>%
  mutate(year = as.numeric(year)) %>% 
  mutate(`Percent Deviation from Trend` = 100*(log(Yield) - log((temp1$coefficients[1] + temp1$coefficients[2]*year)))) %>% 
  ggplot(aes(x = `% Good + Excellent`, y = `Percent Deviation from Trend`)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = purduegold) +
  theme_bw() + 
  #scale_color_manual(values=c("black", purduegold)) + 
  labs(title = "End of Season G + E Crop Condition Ratings and Final Yield")

ggsave("ForecastProduction-PercDeviationVSGEXY.png")
```

# Figures for `09-ForecastingUseof.Rmd` and `10-ForecastingUseofSoy.Rmd`

```{r fig.height = 5}
library(tidyverse)
library(readxl)
library(httr)
library(tidyr)
tmp = 'C:/Users/mindy/Desktop/Documents/PriceAnalysis/PriceAnalysis/assets/feed-grains-yearbook-historical.xlsx'

data_xl <- read_excel(tmp, sheet = "FGYearbookTable04", 
                      skip = 4, 
                      col_names = c("MktYear", "Quarter", "BeginningStocks", "Production", "Imports", 
                                    "TotalSupply", "FoodAlcoholInd", "Seed", "FeedandResidual", 
                                    "TotalDomestic", "Exports", "TotalUse", "EndingStocks")) %>%
  fill(MktYear) %>% 
  # Make columns for more standard date handling
  mutate(month = ifelse(Quarter =="Q1 September-November", "09-01", 
                        ifelse(Quarter == "Q2 December-February", "12-01", 
                               ifelse(Quarter == "Q3 March-May", "03-01", 
                                      ifelse(Quarter == "Q4 June-August", "06-01",
                                             ifelse(Quarter == "MY September-August", "09-01", NA)))))) %>%
  mutate(year = ifelse(Quarter =="Q1 September-November" | Quarter == "Q2 December-February" | Quarter == "MY  September-August", as.numeric(str_sub(MktYear, 1,4)), 
                       as.numeric(str_sub(MktYear, 1,4)) +1 )) %>%
  filter(!is.na(month) & !is.na(year)) %>%
  mutate(date = as.Date(paste0(as.character(year), "-", as.character(month)))) # Now we have a proper date column.

# Make tidy getting ready for ggplot
data_tidy <- data_xl %>% 
  select(c(-MktYear, -month, -year)) %>%
  pivot_longer(c(-date, -Quarter), names_to = "UseCategory", values_to = "Value")

data_tidy
```



Now let's plot the corn dissapearance

```{r, fig.height= 5}
# Plot the Marketing Year Data
data_tidy %>% filter(Quarter != "MY September-August") %>%
  filter(UseCategory %in% c("FoodAlcoholInd", "Seed", "FeedandResidual", "Exports")) %>%
  ggplot(aes(x = date, y = Value, color = UseCategory)) + 
  geom_line() + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Quarterly Use Categories Since 1975")
ggsave(filename = "ForecastingUseof-CornUseCategories.png")

data_tidy %>% filter(Quarter != "MY September-August") %>%
  filter(UseCategory %in% c("FoodAlcoholInd")) %>%
  ggplot(aes(x = date, y = Value)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Quarterly Food Alcohol and Industrial Use Since 1975")
ggsave(filename = "ForecastingUseof-CornUseCategoriesFoodAlcoholInd.png")

data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("FoodAlcoholInd", "TotalUse")) %>%
  pivot_wider(names_from = UseCategory, values_from = Value) %>%
  mutate(ProportionofUse = FoodAlcoholInd/TotalUse) %>%
  ggplot(aes(x = date, y = ProportionofUse)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Marketing Year Food, Alcohol, & Industrial as Proportion of Total Use")
ggsave(filename = "ForecastingUseof-CornUseCategoriesFoodAlcoholIndPropofUse.png")

data_tidy %>% filter(Quarter != "MY September-August") %>%
  filter(UseCategory %in% c("FeedandResidual")) %>%
  ggplot(aes(x = date, y = Value)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Quarterly Feed and Residual Use Since 1975")
ggsave(filename = "ForecastingUseof-CornUseCategoriesFeedResid.png")

data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("FeedandResidual", "TotalUse")) %>%
  pivot_wider(names_from = UseCategory, values_from = Value) %>%
  mutate(ProportionofUse = FeedandResidual/TotalUse) %>%
  ggplot(aes(x = date, y = ProportionofUse)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Marketing Year Feed and Residual as Proportion of Total Use")
ggsave(filename = "ForecastingUseof-CornUseCategoriesFeedandResidPropofUse.png")

data_tidy %>% filter(Quarter != "MY September-August") %>%
  filter(UseCategory %in% c("Exports")) %>%
  ggplot(aes(x = date, y = Value)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Quarterly Exports Since 1975")
ggsave(filename = "ForecastingUseof-CornUseCategoriesExports.png")

data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("Exports", "TotalUse")) %>%
  pivot_wider(names_from = UseCategory, values_from = Value) %>%
  mutate(ProportionofUse = Exports/TotalUse) %>%
  ggplot(aes(x = date, y = ProportionofUse)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Corn Marketing Year Exports as Proportion of Total Use")
ggsave(filename = "ForecastingUseof-CornUseCategoriesExportsPropofUse.png")


```

```{r, fig.height= 5}

data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("FoodAlcoholInd", "Seed", "FeedandResidual", "Exports")) %>%
  ggplot(aes(x = date, y = Value, color = UseCategory)) + 
  geom_line() + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Marketing Year Corn Use Categories Since 1975")
ggsave(filename = "ForecastingUseof-CornUseCategoriesMY.png")

data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("FoodAlcoholInd")) %>%
  ggplot(aes(x = date, y = Value)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Marketing Year Food Alcohol and Industrial Use Since 1975")
ggsave(filename = "ForecastingUseof-CornUseCategoriesFoodAlcoholIndMY.png")

data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("FeedandResidual")) %>%
  ggplot(aes(x = date, y = Value)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Marketing Year Corn Feed and Residual Use Since 1975")
ggsave(filename = "ForecastingUseof-FeedandResidualMY.png")

data_tidy %>% filter(Quarter == "MY September-August") %>%
  filter(UseCategory %in% c("Exports")) %>%
  ggplot(aes(x = date, y = Value)) + 
  geom_line(color = purduegold, size = .75) + 
  theme_bw() +
  labs(x = "", y = "Millions of bushels", title = "Marketing Year Corn Exports Since 1975")
ggsave(filename = "ForecastingUseof-ExportsMY.png")
```

Ethanol

```{r, fig.height= 5}

ethanol <- read_excel("C:/Users/mindy/Desktop/Documents/PriceAnalysis/PriceAnalysis/assets/Table_10.3_Fuel_Ethanol_Overview.xlsx", sheet = 2, skip = 11) 
ethanol <- ethanol[, c(1, 6,12)]
colnames(ethanol) <- c("Year", "Production", "Consumption" )

ethanol %>% pivot_longer(cols = -Year, names_to = "Series", values_to = "MillionGallons") %>% 
  ggplot(aes(x = Year, y = MillionGallons, color = Series)) + 
  geom_line(size = .75) +
  scale_color_manual(values = c(purduegold, "black")) +
  theme_bw() +
  labs(x = "", y = "Millions of gallons", title = "Production and Consumption of Ethanol")

  ggsave(filename = "ForecastingUseof-Ethanol.png")
  

```

# Figures for `12-A-South-American-Production.Rmd`

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width= 6, fig.hight = 4}
library(tidyverse)

# update from FAS Production supply and demand datasets. https://apps.fas.usda.gov/psdonline/app/index.html#/app/downloads 
# all psd_grains_pulses.csv.zip. unzip manually for now. 

grains   <- read.csv('psd_grains_pulses.csv')


start <- 1990

year <-  as.numeric(format(Sys.Date(), format="%Y")) - 1
  
get_top_corn <- grains %>% 
                  filter(Market_Year == year) %>%
                  filter(Commodity_Description == "Corn") %>%
                  filter(Attribute_ID == 28) %>%
                  arrange(desc(Value))

get_top_corn <- get_top_corn[1:5, 'Country_Name']

top_corn <- grains %>%
  filter(Commodity_Description == "Corn") %>%
  filter(Attribute_ID == 28) %>%
  filter(Country_Name %in% get_top_corn) %>%
  filter(Market_Year %in% c(start:year))


c <- ggplot(top_corn, aes(x = Market_Year, y = Value, fill = Country_Name)) + geom_area() + 
      theme_bw() +
      scale_x_continuous(breaks = seq(start, year, 2)) +
      labs(fill = "Top 5 Countries") +
      theme(axis.title.x = element_blank()) +   # Remove x-axis label
      ylab("Production - 1000 MT") + 
      scale_fill_brewer(palette = 'Accent')  +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
      ggtitle("Figure 1: Top World Producers of Corn")  
      
 
c
ggsave(filename = "SouthAmericanProduction_TopProdCorn.png", width= 6, height = 4)

```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width= 6, fig.hight = 4}

get_top_corn <- grains %>% 
                  filter(Market_Year == year) %>%
                  filter(Commodity_Description == "Corn") %>%
                  filter(Attribute_ID == 88) %>%
                  arrange(desc(Value))

get_top_corn <- get_top_corn[1:5, 'Country_Name']

top_corn <- grains %>%
  filter(Commodity_Description == "Corn") %>%
  filter(Attribute_ID == 88) %>%
  filter(Country_Name %in% get_top_corn) %>%
  filter(Market_Year %in% c(start:year))


c2 <- ggplot(top_corn, aes(x = Market_Year, y = Value, fill = Country_Name)) + geom_area() + 
      theme_bw() +
      scale_x_continuous(breaks = seq(start, year, 2)) +
      labs(fill = "Top 5 Countries") +
      theme(axis.title.x = element_blank()) +   # Remove x-axis label
      ylab("Exports - 1000 MT") + 
      scale_fill_brewer(palette = 'Accent')  +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
      ggtitle("Figure 2: Top World Exporters of Corn")  
      
c2
ggsave(filename = "SouthAmericanProduction_TopExporterCorn.png" , width= 6, height = 4)



```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width= 6, fig.hight = 4}
get_top_corn <- grains %>% 
                  filter(Market_Year == year) %>%
                  filter(Commodity_Description == "Corn") %>%
                  filter(Attribute_ID == 88) %>%
                  arrange(desc(Value))

get_top_corn <- get_top_corn[1:10, 'Country_Name']

grains$Production <- grains$Value

grains %>%
              filter(Commodity_Description == "Corn") %>%
              filter(Attribute_ID == 88) %>%
               filter(Country_Name %in% get_top_corn) %>%
              filter(Market_Year == year) %>%
              select(Country_Name, Value) %>%
              arrange(desc(Value)) %>%
  write.csv(file = "corn_exporters.csv")

 
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width= 6, fig.hight = 4}

oilseeds <- read.csv('psd_oilseeds.csv')



get_top_soy <- oilseeds %>% 
                                    filter(Commodity_Code == 2222000) %>%
                  
                  filter(Attribute_ID == 28) %>%
                  filter(Market_Year == as.integer(year-2)) %>%
                  arrange(desc(Value))

get_top_soy <- get_top_soy[1:5, 'Country_Name']

top_soy<- oilseeds %>%
  filter(Commodity_Code == 2222000) %>%
  filter(Attribute_ID == 28) %>%
  filter(Country_Name %in% get_top_soy) %>%
  filter(Market_Year %in% c(start:year))


s <- ggplot(top_soy, aes(x = Market_Year, y = Value, fill = Country_Name)) + geom_area() + 
      theme_bw() +
      scale_x_continuous(breaks = seq(start, year, 2))+
      labs(fill = "Top 5 Countries") +
      theme(axis.title.x = element_blank()) +   # Remove x-axis label
      ylab("Production - 1000 MT") + 
      scale_fill_brewer(palette = 'Accent')  +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
      ggtitle("Figure 3: Top World Producers of Soybeans")  
s
ggsave(filename = "SouthAmericanProd_TopProdSoy.png", width= 6, height = 4)

```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width= 6, fig.hight = 4}






get_top_soy <- oilseeds %>% 
                  filter(Market_Year == year-2) %>%
                  filter(Commodity_Code == 2222000) %>%
                  filter(Attribute_ID == 88) %>%
                  arrange(desc(Value))

get_top_soy <- get_top_soy[1:5, 'Country_Name']

top_soy<- oilseeds %>%
  filter(Commodity_Code == 2222000) %>%
  filter(Attribute_ID == 88) %>%
  filter(Country_Name %in% get_top_soy) %>%
  filter(Market_Year %in% c(start:year))


s2 <- ggplot(top_soy, aes(x = Market_Year, y = Value, fill = Country_Name)) + geom_area() + 
      theme_bw() +
      scale_x_continuous(breaks = seq(start, year, 2))+
      labs(fill = "Top 5 Countries") +
      theme(axis.title.x = element_blank()) +   # Remove x-axis label
      ylab("Exports - 1000 MT") + 
      scale_fill_brewer(palette = 'Accent')  +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
      ggtitle("Figure 4: Top World Exporters of Soybeans")  
s2
ggsave(filename = "SouthAmericanProd_TopExporterSoy.png", height = 4, width = 6)

```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.width= 6, fig.hight = 4}


get_top_soy <- oilseeds %>% 
                  filter(Market_Year == year-2) %>%
                  filter(Commodity_Code == 2222000) %>%
                  filter(Attribute_ID == 88) %>%
                  arrange(desc(Value))

get_top_soy <- get_top_soy[1:10, 'Country_Name']

oilseeds$Production <- oilseeds$Value

oilseeds %>%
              filter(Commodity_Code == 2222000) %>%
              filter(Attribute_ID == 88) %>%
               filter(Country_Name %in% get_top_soy) %>%
              filter(Market_Year == year) %>%
              select(Country_Name, Value) %>%
              arrange(desc(Value)) %>%
  write.csv(file = "soy_exporters.csv")



```

# Figures for `11-EndingStocksand.Rmd`

```{r}
tmp <- tempfile(fileext = ".xls")
library(httr)

tmp = 'C:/Users/mindy/Desktop/Documents/PriceAnalysis/PriceAnalysis/assets/feed-grains-yearbook-historical.xlsx'

data_xl <- read_excel(tmp, sheet = "FGYearbookTable04", 
                      skip = 4, 
                      col_names = c("MktYear", "Quarter", "BeginningStocks", "Production", "Imports", 
                                    "TotalSupply", "FoodAlcoholInd", "Seed", "FeedandResidual", 
                                    "TotalDomestic", "Exports", "TotalUse", "EndingStocks")) %>%
  fill(MktYear) %>% 
  # Make columns for more standard date handling
  mutate(month = ifelse(Quarter =="Q1 September-November", "09-01", 
                        ifelse(Quarter == "Q2 December-February", "12-01", 
                               ifelse(Quarter == "Q3 March-May", "03-01", 
                                      ifelse(Quarter == "Q4 June-August", "06-01",
                                             ifelse(Quarter == "MY September-August", "09-01", NA)))))) %>%
  mutate(year = ifelse(Quarter =="Q1 September-November" | Quarter == "Q2 December-February" | Quarter == "MY  September-August", as.numeric(str_sub(MktYear, 1,4)), 
                       as.numeric(str_sub(MktYear, 1,4)) +1 )) %>%
  filter(!is.na(month) & !is.na(year)) %>%
  mutate(date = as.Date(paste0(as.character(year), "-", as.character(month)))) # Now we have a proper date column.

# # Make tidy getting ready for ggplot
# data_tidy <- data_xl %>% 
#   select(c(-MktYear, -month, -year)) %>%
#   pivot_longer(c(-date, -Quarter), names_to = "UseCategory", values_to = "Value")
# 
# data_tidy
```

```{r}



data_fg <- data_xl %>% 
  filter(Quarter == "MY September-August") %>% 
  select(c(date, TotalUse, EndingStocks)) %>%
  mutate(date = year(date)) %>% 
  mutate(StockstoUse = EndingStocks/TotalUse*100)

key = 'BD652A51-1B61-3442-B80A-9B4030E2F382'


# apikey <- "YOUR API KEY"
nassqs_auth(key)

params1 <- list(commodity_desc = "CORN", 
                year__GE = 1975, 
                agg_level_desc = "NATIONAL", 
                statisticcat_desc = "PRICE RECEIVED",
                #prodn_practice_desc = "ALL PRODUCTION PRACTICES",
                #util_practice_desc = c("GRAIN", "ALL UTILIZATION PRACTICES"),
                reference_period_desc = "Marketing Year",
                freq_desc = "ANNUAL",
                unit_desc = "$ / BU"
                #prodn_practice_desc = "NON-IRRIGATED"
                ) # GE means "greater than or equal to"
data1   <- nassqs(params1)
data1$Value <- as.numeric(gsub(",", "", as.character(data1$Value)))

data1 <- data1 %>% mutate(date = as.double(year))

data1 <- inner_join(data_fg, data1 %>% select(date, Value), by = "date") %>% select(date, StockstoUse, Value)
colnames(data1) <- c("date", "StockstoUse", "PricesRecieved")
data1 <- data1 %>% mutate(PricesRecieved2 = 10*PricesRecieved)
```

```{r}

ppp<- data1 %>% pivot_longer(cols = c(StockstoUse, PricesRecieved2), 
                       names_to = "Series", 
                       values_to = "value") %>%
  ggplot(aes(x = date, y = value, color = Series)) + 
  geom_line(size = 0.75) +
  scale_color_manual(values = c(purduegold, "black")) + 
  theme_bw() + 
  scale_y_continuous(# Features of the first axis
    name = "Stocks-to-Use %",
    # Add a second axis and specify its features
    sec.axis = sec_axis( trans=~./10, name="$/bu Prices Received")) +
  labs(title = "Corn Stocks-to-Use and Prices Received by Farmers", 
       x = "", 
       y = "$/bu and Stocks/TotalUse*100",
       caption = "1975-2020 Marketing Years (2020 Estimated as of 2021-03-23)")
ppp
ggsave(ppp, filename = "EndingStocksand-StocksUsePrices.png" , width= 6, height = 4)

```

```{r}

data1 %>% 
  ggplot(aes(x = StockstoUse, y = PricesRecieved)) + 
  geom_point(size = 1.5) +
  scale_color_manual(values = c(purduegold, "black")) + 
  theme_bw() + 
  labs(title = "Corn Stocks-to-Use and Prices Received by Farmers", 
       x = "Stocks/TotalUse*100", 
       y = "$/bu",
       caption = "1975-2025 Marketing Years (2025 Estimated as of 2025-12-20)")

ggsave(filename = "EndingStocksand-StocksUsePrices2.png" , width= 6, height = 4)

```

```{r}

data1 %>% mutate(BeforeAfter = case_when(date < 2006 ~ "Before",
                           date >= 2006 ~ "After")) %>%
  ggplot(aes(x = StockstoUse, y = PricesRecieved, color = BeforeAfter)) + 
  geom_point(size = 1.5) +
  geom_smooth(method = lm, formula = y ~ 1 + log(x), se = FALSE) +
  scale_color_manual(values = c(purduegold, "black")) + 
  theme_bw() + 
  labs(title = "Corn Stocks-to-Use and Prices Received by Farmers, Before and After 2006", 
       x = "Stocks/TotalUse*100", 
       y = "$/bu",
       caption = "1975-2025 Marketing Years (2025 Estimated as of 2025-12-20)")

ggsave(filename = "EndingStocksand-StocksUsePrices3.png" , width= 6, height = 4)

```

# Figures for `13-EthanolMarketsand.Rmd`

```{r}
data <- read_excel('d1-10ethanolprofitabilit.xlsx', 
           sheet = "Returns per Bu.", 
           range = "A8:J1000",
           col_types = c("date", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric", "numeric")) %>% 
  select(c(1, 4, 6, 8, 10))

colnames(data) <- c("Date", "Ethanol", "DDG", "Corn", "NaturalGas")

```

```{r}

data %>% mutate(DDG = DDG/2000*17) %>% 
  pivot_longer(cols = c(Ethanol, DDG, Corn, NaturalGas), names_to = "Commodity", values_to = "Prices") %>% 
  filter(Commodity == "Corn" | Commodity == "DDG") %>% 
  ggplot(aes(x = Date, y = Prices, color = Commodity)) +
  geom_line(size = .75) + 
  theme_bw() + 
  scale_color_manual(values = c(purduegold, "black")) +
  labs(title = "Corn and DDG Prices, 2005-2021")

ggsave("CornDDGPrices.png", width = 6, height = 4)

```

```{r}

data %>% 
   mutate(Ratio = DDG/Corn) %>%
  ggplot(aes(x = Date, y = Ratio)) +
  geom_line(size = .75, color = purduegold) + 
  theme_bw() + 
  #scale_color_manual(values = c(purduegold, "black")) +
  labs(title = "DDG/Corn Price Ratio, 2005-2021")

ggsave("DDGCORNRatio.png", width = 6, height = 4)

```

```{r}

data %>% 
   mutate(Crush = Ethanol*2.8 + DDG*17/2000 - Corn) %>%
  ggplot(aes(x = Date, y = Crush)) +
  geom_line(size = .75, color = purduegold) + 
  theme_bw() +
  lims(y = c(-1,8)) +
  #scale_color_manual(values = c(purduegold, "black")) +
  labs(title = "Ethanol Crush Spread, 2005-2021")

ggsave("ECrushSpread.png" , width = 6, height = 4)

```

# Figures for `12-SoybeanCrush.Rmd`

```{r}
S <- read.csv(url(paste0("http://ondemand.websol.barchart.com/getHistory.csv?apikey=", 
                         KEY, 
                         "&symbol=ZS*1&type=dailyNearest&backAdjust=false&startDate=20060201"))) %>%
   mutate(tradingDay = as.Date(tradingDay))

BO <- read.csv(url(paste0("http://ondemand.websol.barchart.com/getHistory.csv?apikey=", 
                          KEY, 
                          "&symbol=ZL*1&type=dailyNearest&backAdjust=false&startDate=20060201"))) %>%
   mutate(tradingDay = as.Date(tradingDay))

BM <- read.csv(url(paste0("http://ondemand.websol.barchart.com/getHistory.csv?apikey=", 
                          KEY, 
                          "&symbol=ZM*1&type=dailyNearest&backAdjust=false&startDate=20060201"))) %>%
   mutate(tradingDay = as.Date(tradingDay))

#S %>% mutate(spread = )

BM %>% ggplot(aes(x = tradingDay, y = close)) + geom_line()


df = data.frame(Date = S$tradingDay, y = BO$close*11/100+ BM$close*.022- S$close/100 )

pp <- df %>% ggplot(aes(x = Date, y = y)) + 
  geom_line() +
  theme_bw() + 
  labs(y = "Soybean Crush Spread, $/bu")
pp
ggsave("SoyCrushSpread.png" , width = 6, height = 4)
```
